---
title: "NY AIRBNB ANALYSIS"
author: "MEGHANA ATLURI"
format: pdf
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  tidy = TRUE,
  comment = "#>",
  fig.align = 'center',
  echo = TRUE
)
```

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(janitor)
library(lubridate)
```

### [I . Reading CSV File]{.underline}

```{r}
#Impoting File
ABNB <- read_csv("AB_NYC_2019.csv")
```

So we have a total of 48895 numbers of rows and 16 numbers of columns (features) in our dataset.

### [II. UNDERSTANDING OUR DATA]{.underline}

```{r}
#dimensions of dataset
dim(ABNB)
```

So we have a total of 48895 numbers of rows and 16 numbers of columns (features) in our dataset.

```{r}
# to see type associated with each column, and the first few values of each column of the data frame.
glimpse(ABNB)
```

```{r}
head(ABNB)
```

```{r}
#column names in our dataset
colnames(ABNB)
```

### [III.DATA CLEANING]{.underline}

**DROP UNNECESSARY COLUMNS**

```{r}
#Dropping columns by index
ABNB <- select(ABNB, -c(1,2,12,13))
```

Dropping columns listing id, name, minimum nights, number of reviews, last review from dataset.

**HANDLING NA VALUES**

```{r}
#Checking number of NA values in each column
count_na <- function(x) {
  sum(is.na(x))
}
apply(ABNB, MARGIN = 2, FUN = count_na)
```

```{r}
# Replace NA with "UNKNOWN" in the hostname column
ABNB$host_name[is.na(ABNB$host_name)] <- "Unknown"
```

```{r}
# Filling missing values; assuming missing reviews indicate no reviews yet
ABNB$reviews_per_month <- replace(ABNB$reviews_per_month, is.na(ABNB$reviews_per_month), 0)
```

**CLEANING AND RENAMING COLUMNS**

```{r}
#cleaning column names and renaming them
ABNB <- ABNB |>
  clean_names() |>
  rename( borough_name = neighbourhood_group,
         neighbourhood_name = neighbourhood , 
        host_listing_count = calculated_host_listings_count,
        availability_per_year = availability_365) 

```

**AS FACTORS**

```{r}
#CONVERTING BOROUGH NAME TO FACTOR
 ABNB$borough_name = as.factor(ABNB$borough_name)
#CONVERTING ROOM TYPE TO FACTOR
 ABNB$room_type = as.factor(ABNB$room_type)
```

### [IV. DATA ANALYSIS]{.underline}

#### **Analyze borough in NYC**

```{r}
#analysis of borough name, reorder the factor
ggplot(ABNB, aes(x = fct_infreq(borough_name))) +
geom_bar(fill="forestgreen") + 
  xlab("Borough Name")+
  ylab("Number of Listings") +
   ggtitle("Number of Airbnb Listings by Borough")
```

This plot helps identify which boroughs are most active on Airbnb - Manhattan and Brooklyn have the highest listing counts , and Staten Island has least number of listings in NYC.

#### **Analyze room_type in NYC**

```{r}
#analysis of borough name, reorder the factor
ggplot(ABNB, aes(x = fct_rev(fct_infreq(room_type)))) +
geom_bar(fill="steelblue" ) + 
  xlab("Room Type")+
    ylab("Number of Listings") +
  ggtitle("Number of Airbnb Listings by Room Type")
```

This plot reveal which types of accommodations are most common in NYC’s Airbnb market “Entire home/apt” is often the most listed type, followed by “Private room.”

#### **Count of room types in different boroughs in NYC**

```{r}
#analyze room types in different borough
ggplot(data = ABNB) +
geom_bar(aes(x = room_type, fill = borough_name),position = "dodge") + coord_flip() +
  xlab("Room Type") +
  ylab("Number of Listings") +
  ggtitle("Distribution of Room Types Across NYC Boroughs")
```

This plot highlights how boroughs differ in their room type offerings. For instance, Manhattan has more entire home/apt, while Brooklyn lean toward private rooms. It’s useful for comparing room type trends by location.

#### **Coordinates of Airbnb Room type available across NYC**

```{r}
#analyze coordinates for different room types available
ggplot(data = ABNB, aes(x = latitude, y = longitude, color = room_type)) +
geom_point(size = 1, alpha = 0.6) +
  xlab("Latitude")+
  ylab("Longitude")+
  ggtitle("Geographic Distribution of Airbnb Room Types in NYC") +
  scale_color_brewer(palette = "Dark2")
```

#### **Comparing price across all borough for different room types in NYC 2019.**

```{r}
#price is distributed for different room type in each boroughs across nyc 
ggplot(data = ABNB, aes(x = borough_name, y = price,fill=room_type)) +
  geom_boxplot()+
  coord_cartesian(ylim = c(0, 500)) +
  xlab("Borough Name")+
  ylab("Price (in dollars)")+
    ggtitle("Room Type Price Distribution Across NYC Boroughs")+
    scale_fill_brewer(palette = "Pastel1")
```

-   Entire home/apt has the highest median price over other room types, followed by Private room and shared room, as expected

-   Entire homes in Manhattan have the highest median prices. Shared rooms and private rooms are more affordable, especially in boroughs like the Bronx

-   Price variability also higher in tourist-heavy boroughs. (Manhattan,Brooklyn,Queens)

```{r}
#using facets to compare price for rooms in different boroughs in NYC
ggplot(data = ABNB, aes(x = room_type, y = price, fill= room_type)) +
geom_violin() +
coord_cartesian(ylim = c(0, 500)) +
  xlab("Room Type")+
  ylab("Price (in dollars)") +
facet_wrap(vars(borough_name), ncol=1 )
```

#### **Room_types and their relation with availability in different Boroughs**

```{r}
ggplot(ABNB,aes(room_type,availability_per_year))+
  geom_boxplot(aes(fill=borough_name))+
  xlab("Room Type") +
  ylab("Availability in a year") +
  ggtitle("Availability of rooms in different boroughs in NYC")
```

Brooklyn and Manhattan have the lowest availability i.e it has highest demand. Shared rooms tend to have more availability meaning less demand.

```{r}
ABNB |>
group_by(borough_name,room_type)|>
  summarize(
    count=n() , 
    avg.availability= mean(availability_per_year) 
  ) |>
arrange(desc(avg.availability))
```

#### **Top seen hosts with listings in NYC**

```{r}
ABNB3 <- ABNB |>
  select(host_id, host_name, host_listing_count) |>
  distinct() |>
  arrange(desc(host_listing_count)) |>
  slice_head(n = 10)


#plotting Top10 hosts with listing in NYC
ggplot(ABNB3, aes(x = reorder(host_name, host_listing_count), y = host_listing_count)) +
  geom_col(fill = "pink") +
  geom_text(aes(label = host_listing_count)) +
  coord_flip() +
  xlab("Host Name") +
  ylab("Listing Count") +
  ggtitle("Top 10 Airbnb Hosts in NYC by Listings") 
```

Top Hosts with listings in NYC in year 2019 is Sonder , Blueground , Kara , Kazuya and so on...

```{r}
# Select relevant columns
reviews <- select(ABNB, host_id, reviews_per_month)

# Join with top 10 hosts table
final <- inner_join(ABNB3, reviews, by = "host_id")

# Group and summarize
ABNB3B <- final |>
  group_by(host_name) |>
  summarize(
    avg_review = mean(reviews_per_month, na.rm = TRUE),
    avg_count_listing = mean(host_listing_count, na.rm = TRUE)
  )
# Correlation between avg listings and avg reviews
cor_value <- cor(ABNB3B$avg_review, ABNB3B$avg_count_listing, use = "complete.obs")
round(cor_value, 3)

```

This result suggests a strong positive correlation (≈ 0.78) between the average number of reviews per host and their average number of listings. Hosts who manage more listings tend to receive more reviews.

#### **Which Neighbourhood and Room type has highest reviews in 2019**

```{r}
# Summarize reviews by borough and room type
ABNB4 <- ABNB |>
  select(borough_name,room_type , reviews_per_month) |>
group_by(borough_name,room_type)|>
summarize(
  count=n(),
  avg_review = mean(reviews_per_month)
)

ggplot(ABNB4, aes(x = room_type, y = avg_review, fill = borough_name)) +
  geom_col(position = "dodge") +
  xlab("Room Type") +
  ylab("Average Reviews per Month") +
  ggtitle("Average Monthly Reviews by Room Type and Borough") +
  theme_minimal()
```

-   In most boroughs "Entire apt/home" is high rated followed by "Private" and "Shared room".

-   Shared rooms are least reviewed overall, especially in Staten Island and Brooklyn, indicating lower guest preference or listing volume.

### [V. MODELING]{.underline}

```{r}
# Linear regression model
model <- lm(price ~ borough_name + room_type + availability_per_year + reviews_per_month + minimum_nights, data = ABNB)

# Model summary
summary(model)

```

```{r}
# Getting model diagnostics
par(mfrow = c(2, 2))
plot(model)
```

The graphs above clearly tell that its nonlinear, residuals not normally distributed, and indicates heteroscedasticity

So, next step is **log-transformed model.**

```{r}
# Clean the dataset: remove rows with NA or non-positive price
ABNB <- ABNB[!is.na(ABNB$price) & ABNB$price > 0, ]

# Log-transform the price again to overwrite any bad values
ABNB$log_price <- log(ABNB$price)

# Fit the linear model
log_model <- lm(log_price ~ borough_name + room_type + availability_per_year + reviews_per_month + minimum_nights, data = ABNB)

# View model summary
summary(log_model)

```

```{r}
# Diagnostics plots
par(mfrow = c(2, 2))
plot(log_model)
```

```{r}
summary(model)$r.squared
summary(log_model)$r.squared
```

```{r}
AIC(model, log_model)
BIC(model, log_model)
```

From R-squared , AIC , BIC it is clear that log_model is better model than base model.
